---
title: "PresSlideShow"
output: ioslides_presentation
date: "2023-07-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(dplyr)
library(zipcodeR)
library(ggplot2)
```

```{r}
snap_master %>% 
  filter(`COLLEGE CODE` == "SS") %>% 
  aggregate(ret_fall ~ `MAJOR CODE 1`,data = .,mean) %>%
  ggplot(aes(x=`MAJOR CODE 1`,y=ret_fall, group= `MAJOR CODE 1`))+
  geom_col()+
  labs(title = "Rentention by Major in SS",
       x = "Major",
       y = "Rentention Rate")+
  theme_light()
```


```{r}
snap_master %>% 
  ggplot(aes(`REPORT ETHNICITY`))+
  geom_bar(aes(y = ..prop..,group = 1))+
  labs(title= "Ethnicity of XU Students",
       x= "Ethnicity",
       y= "Proportion of Students")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
---
```{r}
snap_master %>% 
aggregate(ret_fall ~ `COMMUTER DORM`+`COHORT TERM`+`REPORT ETHNICITY`,data = .,mean) %>%
ggplot(aes(x=`COHORT TERM`,y=ret_fall,color=`COMMUTER DORM`, group=`COMMUTER DORM`))+
geom_point()+
geom_line()+
facet_wrap(~`REPORT ETHNICITY`) +
ylab("Retention to Fall")+
xlab("Cohort Term")+
theme_bw()
```
---

```{r}
snap_master %>% 
  filter(RELIGION %in% c("Roman Catholic", "Refused", "Other Christian", "No religion", "Baptist")) %>% 
  aggregate(ret_fall ~ RELIGION+`COHORT TERM`,data = .,mean) %>%
  ggplot(aes(x=`COHORT TERM`,y=ret_fall,color=RELIGION, group=RELIGION))+
  geom_point()+
  geom_line()+
  ylab("Retention to Fall")+
  xlab("Cohort Term")+
  theme_bw()
```
---
```{r}
snap_master %>% 
  filter(RELIGION %in% c("Roman Catholic", "Refused", "Other Christian", "No religion", "Baptist")) %>% 
  aggregate(ret_spring ~ RELIGION+`COHORT TERM`,data = .,mean) %>%
  ggplot(aes(x=`COHORT TERM`,y=ret_spring,color=RELIGION, group=RELIGION))+
  geom_point()+
  geom_line()+
  ylab("Retention to Spring")+
  xlab("Cohort Term")+
  theme_bw()
```

---
```{r}
snap_master %>% 
  filter(`HOME STATE` %in% c("OH", "KY", "IN", "IL", "MI", "NY" )) %>% 
aggregate(ret_fall ~ `HOME STATE`+`COHORT TERM`,data = .,mean) %>%
ggplot(aes(x=`COHORT TERM`,y=ret_fall,color=`HOME STATE`, group=`HOME STATE`))+
geom_point()+
  geom_line()+
ylab("Retention to Fall")+
xlab("Cohort Term")+
theme_bw()
```
---
```{R}
snap_master %>% 
  filter(`HOME STATE` %in% c("OH", "KY", "IN", "IL", "MI", "NY" )) %>% 
aggregate(ret_spring ~ `HOME STATE`+`COHORT TERM`,data = .,mean) %>%
ggplot(aes(x=`COHORT TERM`,y=ret_spring,color=`HOME STATE`, group=`HOME STATE`))+
geom_point()+
  geom_line()+
ylab("Retention to Spring")+
xlab("Cohort Term")+
theme_bw()

```
---


```{r}
snap_master %>% 
  ggplot(aes(x=`HS GPA`,color= ret_fall, group= ret_fall))+
  geom_histogram ()+
  ylab("Retention to Fall")+
  xlab("Cohort Term")+
  theme_bw()
```
---


GPA Histograms
```{r}
snap_master %>%
  ggplot(aes(GPA.y, fill = ret_fall))+
  geom_histogram(aes(), alpha = 0.5, position = "identity") +
  labs(title = "Xavier Students GPA",
       x = "GPA",
       y = "Count")+
  scale_fill_manual(values = c("black", "pink"))+
  scale_color_manual(values = c("black","pink"))
```
--- 
```{r}
snap_master %>%
  ggplot(aes(`HS GPA`, fill = ret_fall))+
  geom_histogram(aes(), alpha = 0.5, position = "identity") +
  labs(title = "Xavier Students High School GPA",
       x = "GPA",
       y = "Count")+
  scale_fill_manual(values = c("black","pink"))+
  scale_color_manual(values = c("black","pink"))
```
---
Zip code distance
```{r}
snap_master <- snap_master %>%
mutate(recoded_zip_code = substr(`HOME ZIP`, 1, 5))
snap_master <- snap_master %>%
  mutate(Distance = zip_distance(45207, `recoded_zip_code`)$distance)
snap_master %>% 
  ggplot(aes(Distance, GPA.y, color = ret_fall))+
  geom_point()+
  geom_smooth()
```
---
Zip code distance

```{r}
snap_master %>% 
  ggplot(aes(log(Distance), `INST AID AMT`, color = ret_fall))+
  geom_point()+
  geom_smooth()
```

---
Black/White/Commuter/Resident
```{r}

white_resident_ret <- snap_master %>%
  filter(`REPORT ETHNICITY` == "White", `COMMUTER DORM` == "R")

black_commuter_ret <- snap_master %>%
  filter(`REPORT ETHNICITY` == "Black or African-American", `COMMUTER DORM` == "C")

black_resident_ret <- snap_master %>%
  filter(`REPORT ETHNICITY` == "Black or African-American", `COMMUTER DORM` == "R")

white_commuter_ret <- snap_master %>%
  filter(`REPORT ETHNICITY` == "White", `COMMUTER DORM` == "C")

white_resident_prop <- white_resident_ret %>%
  summarise(Yes = mean(ret_fall == TRUE))

black_commuter_prop <- black_commuter_ret %>%
  summarise(Yes = mean(ret_fall == TRUE))

black_resident_prop <- black_resident_ret %>%
  summarise(Yes = mean(ret_fall == TRUE))

white_commuter_prop <- white_commuter_ret %>%
  summarise(Yes = mean(ret_fall == TRUE))

retention_data <- data.frame(
  Group = c("White Resident", "Black Commuter", "Black Resident", "White Commuter"),
  Retained = c(white_resident_prop$Yes, black_commuter_prop$Yes, black_resident_prop$Yes, white_commuter_prop$Yes)
)

ggplot(retention_data, aes(x = Group)) +
  geom_bar(aes(y = Retained), stat = "identity", fill = "blue", width = 0.4) +
  labs(title = "Retention Rates by Student Type",
       x = "Student Type",
       y = "Proportion of Students Retained",
       fill = "Retention Status") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme_minimal()
```
---
```{r}
library(dplyr)
library(ggplot2)


# Black/White/Commuter/Resident/GPA>3/GPA<3
white_resident_gpa_gt3 <- snap_master %>%
  filter(`REPORT ETHNICITY` == "White", `COMMUTER DORM` == "R", `HS GPA` > 3)

white_resident_gpa_lt3 <- snap_master %>%
  filter(`REPORT ETHNICITY` == "White", `COMMUTER DORM` == "R", `HS GPA` < 3)

white_commuter_gpa_gt3 <- snap_master %>%
  filter(`REPORT ETHNICITY` == "White", `COMMUTER DORM` == "C", `HS GPA` > 3)

white_commuter_gpa_lt3 <- snap_master %>%
  filter(`REPORT ETHNICITY` == "White", `COMMUTER DORM` == "C", `HS GPA` < 3)

black_resident_gpa_gt3 <- snap_master %>%
  filter(`REPORT ETHNICITY` == "Black or African-American", `COMMUTER DORM` == "R", `HS GPA` > 3)

black_resident_gpa_lt3 <- snap_master %>%
  filter(`REPORT ETHNICITY` == "Black or African-American", `COMMUTER DORM` == "R", `HS GPA` < 3)

black_commuter_gpa_gt3 <- snap_master %>%
  filter(`REPORT ETHNICITY` == "Black or African-American", `COMMUTER DORM` == "C", `HS GPA` > 3)

black_commuter_gpa_lt3 <- snap_master %>%
  filter(`REPORT ETHNICITY` == "Black or African-American", `COMMUTER DORM` == "C", `HS GPA` < 3)

#
white_resident_gt3_prop <- white_resident_gpa_gt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))

white_resident_lt3_prop <- white_resident_gpa_lt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))

white_commuter_gt3_prop <- white_commuter_gpa_gt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))

white_commuter_lt3_prop <- white_commuter_gpa_lt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))

black_resident_gt3_prop <- black_resident_gpa_gt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))

black_resident_lt3_prop <- black_resident_gpa_lt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))

black_commuter_gt3_prop <- black_commuter_gpa_gt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))

black_commuter_lt3_prop <- black_commuter_gpa_lt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))

#
retention_data <- data.frame(
  Group = c("White Commuter GPA < 3", "Black Commuter GPA < 3",
            "Black Resident GPA < 3", "Black Resident GPA > 3",
            "White Resident GPA < 3", "Black Commuter GPA > 3",
            "White Commuter GPA > 3", "White Resident GPA > 3"),
  Retained = c(white_commuter_lt3_prop$Yes, black_commuter_lt3_prop$Yes,
               black_resident_lt3_prop$Yes, black_resident_gt3_prop$Yes,
               white_resident_lt3_prop$Yes, black_commuter_gt3_prop$Yes,
               white_commuter_gt3_prop$Yes, white_resident_gt3_prop$Yes)
)

#
retention_data$Group <- factor(retention_data$Group, levels = c(
  "White Commuter GPA < 3", "Black Commuter GPA < 3",
  "Black Resident GPA < 3", "Black Resident GPA > 3",
  "White Resident GPA < 3", "Black Commuter GPA > 3",
  "White Commuter GPA > 3", "White Resident GPA > 3"
))

ggplot(retention_data, aes(x = Group)) +
  geom_bar(aes(y = Retained), stat = "identity", fill = "blue", width = 0.4) +
  labs(title = "Retention Rates by Student Type and GPA",
       x = "Student Type and GPA",
       y = "Proportion of Students Retained",
       fill = "Retention Status") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))  # Rotate x-axis labels to be vertical
```
***
Commuter/Resident/GPA<3/GPA>3
```{r}

commuter_gt3 <- snap_master %>%
  filter(`COMMUTER DORM` == "C", `HS GPA` > 3)

commuter_lt3 <- snap_master %>%
  filter(`COMMUTER DORM` == "C", `HS GPA` < 3)

resident_gt3 <- snap_master %>%
  filter(`COMMUTER DORM` == "R", `HS GPA` > 3)

resident_lt3 <- snap_master %>%
  filter(`COMMUTER DORM` == "R", `HS GPA` < 3)


commuter_gt3_prop <- commuter_gt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))

commuter_lt3_prop <- commuter_lt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))

resident_gt3_prop <- resident_gt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))

resident_lt3_prop <- resident_lt3 %>%
  summarise(Yes = mean(ret_fall == TRUE))


retention_data <- data.frame(
  Group = c("Commuter GPA > 3", "Commuter GPA < 3",
            "Resident GPA > 3", "Resident GPA < 3"),
  Retained = c(commuter_gt3_prop$Yes, commuter_lt3_prop$Yes,
               resident_gt3_prop$Yes, resident_lt3_prop$Yes)
)


retention_data$Group <- factor(retention_data$Group, levels = c(
  "Commuter GPA > 3", "Commuter GPA < 3",
  "Resident GPA > 3", "Resident GPA < 3"
))

ggplot(retention_data, aes(x = Group)) +
  geom_bar(aes(y = Retained), stat = "identity", fill = "blue", width = 0.4) +
  labs(title = "Retention Rates by Student Type and GPA",
       x = "Student Type and GPA",
       y = "Proportion of Students Retained",
       fill = "Retention Status") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))  # Rotate x-axis labels to be vertical
```
---
Black/White/GPA>3/GPA<3
```{r}

# 
race_gt3 <- snap_master %>%
  filter(`GPA.y` > 3)

race_lt3 <- snap_master %>%
  filter(`GPA.y` < 3)

#
white_gt3_prop <- race_gt3 %>%
  filter(`REPORT ETHNICITY` == "White") %>%
  summarise(Yes = mean(ret_fall == TRUE))

white_lt3_prop <- race_lt3 %>%
  filter(`REPORT ETHNICITY` == "White") %>%
  summarise(Yes = mean(ret_fall == TRUE))

black_gt3_prop <- race_gt3 %>%
  filter(`REPORT ETHNICITY` == "Black or African-American") %>%
  summarise(Yes = mean(ret_fall == TRUE))

black_lt3_prop <- race_lt3 %>%
  filter(`REPORT ETHNICITY` == "Black or African-American") %>%
  summarise(Yes = mean(ret_fall == TRUE))

#
retention_data <- data.frame(
  Group = c("White GPA > 3", "White GPA < 3",
            "Black GPA > 3", "Black GPA < 3"),
  Retained = c(white_gt3_prop$Yes, white_lt3_prop$Yes,
               black_gt3_prop$Yes, black_lt3_prop$Yes)
)

#
retention_data$Group <- factor(retention_data$Group, levels = c(
  "White GPA > 3", "White GPA < 3",
  "Black GPA > 3", "Black GPA < 3"
))

ggplot(retention_data, aes(x = Group)) +
  geom_bar(aes(y = Retained), stat = "identity", fill = "blue", width = 0.4) +
  labs(title = "Retention Rates by Race and GPA",
       x = "Race and GPA",
       y = "Proportion of Students Retained",
       fill = "Retention Status") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))  # Rotate x-axis labels to be vertical
```